Generating the Blueprints of the Java Ecosystem

### 2015-02-28 msr.tex
changes in Section 2

The dataset construction process is illustrated in Figure~\ref{fig:dataset-construction}. In general, it follows the same methodology presented in reference \cite{MKLGS14}. Initially, a snapshot of the Maven repository was downloaded locally. The repository contains various projects that contain several versions. In our experiment, we only used Java projects. For every Java project
all versions were filtered out and only the latest was kept. The process to identify them was the following; Every project consists of two {\sc jar} files, one that contains a compiled version of the project and one that contains the source code. The projects that did not had both binary and source jars, were excluded from the experiment. When the source {\sc jar} file was downloaded, it was scanned for Java source files. If source files were present in the archive, then the project was flagged as Java project and considered valid.

When the selection process was finished, we created a series of processing tasks based on the containing all the selected {\sc jar}s, and added them to a task queue mechanism. Then we executed a number of workers that checked out tasks from the queue, applied the static analysis tools on each {\sc jar} and stored the results to the data repository (a {\sc m}y{\sc sql} database system). The tools that were applied on every {\sc jar} were the following. The workers were written in Python. In addition, there were several shell scripts (bash) that performed several checks and sanitisation tasks.

\begin{itemize}
  \item \textbf{CKJM-ext} \textit{(ckjm)}~\cite{Spi05g} that calculates many software metrics, including the \textit{Chidamber and Kemmerer} set of object-oriented metrics \cite{CHKE94}. The version of the tool that we used for this experiment can be found in GitHub~\cite{CKJM}.

  \item \textbf{JDepend} \textit{(jdep)}~\cite{JDEPEND} is a tool that analyses {\sc jar} files that contain compiled Java classes and calculates a series of design metrics.

  \item \textbf{CLMT} \textit{(clmt)}~\cite{SGKL09} stands for \textit{cross-language metric tool} and analyses the source code of several languages in order to calculate a series of size metrics.
\end{itemize}

@@

Table \ref{tbl:oss-size-metrics} presents several size metrics for our dataset. It includes 11,365 projects, with more than 74 million lines of code and almost 33 million unique measurements.

Table~\ref{tbl:class-selected-metrics}, \ref{tbl:method-selected-metrics}, \ref{tbl:package-selected-metrics}, and \ref{tbl:size-selected-metrics} present the key metrics that are calculated and stored in the dataset. Each tool focuses on a different aspect of a software system; \textit{ckjm} focuses on object-oriented design metrics, \textit{jdepend} on package design, while \textit{clmt} focuses on program size metrics. Several metrics are calculated by more than one tool, like \textit{cyclomatic complexity}, which is calculated by both \textit{clmt} and \textit{ckjm}. Both calculations are available in the database.

A series of newly introduced metrics were also stored in the database that count the usage of specific {\sc dsl} application libraries in a software project. These are included in the Table~\ref{tbl:size-selected-metrics}. The methodology to identify and calculate the {\sc dsl} metrics is the following: A set of standard {\sc dsl} application libraries was identified, and the source was scanned for specific \textit{import} statements e.g. \textit{java.util.regex}, which indicated that the standard package that implements regular expressions was used, thus regular expressions were used in the project. Build files or other resources that may contain {\sc dsl}s were not included. One final assumption was also made; if {\sc xpath} or {\sc xslt} were found in the source code, then the project would be marked that it also uses {\sc xml}, since those two languages are used for query and transformations on {\sc xml} {\sc dom} trees. Table \ref{tbl:dsl-list} lists the selected {\sc dsl}s application libraries. Note that we focused on libraries that were  included as part of the Java {\sc sdk}.

@@

In Section~\ref{sec:dsl}, we describe an experiment based on these metrics that measures the popularity of {\sc dsl} usage for the dataset and therefore for the Java ecosystem in general.

>>>>

The dataset construction process is illustrated in Figure~\ref{fig:dataset-construction}. In general, it follows the same methodology presented in reference \cite{MKLGS14}.

Initially, a snapshot of the Maven repository was downloaded locally. The repository contains various projects that contain several versions. In our experiment, we only used Java projects. For every Java project all versions were filtered out and only the latest was kept. The process to identify them was the following; Every project consists of two {\sc jar} files, one that contains a compiled version of the project and one that contains the source code. The projects that did not had both binary and source {\sc jar}s, were excluded from the experiment. When the source {\sc jar} file was downloaded, it was scanned for Java source files. If the source files were present in the archive, then the project was flagged as a valid Java project.

When the selection process was finished, we created a series of processing tasks based on the selected {\sc jar}s, and added them to a task queue mechanism. Then we executed a number of workers that checked out tasks from the queue, applied the static analysis tools on each {\sc jar} and stored the results to the data repository (a {\sc m}y{\sc sql} database system). The workers were written in Python. In addition, there were several shell scripts (bash) that performed several checks and sanitisation tasks (e.g. checking for malformed results). The tools that were applied on every {\sc jar} were the following:

\begin{itemize}
  \item \textbf{CKJM-ext} \textit{(ckjm)}~\cite{Spi05g}; which calculates many software metrics, including the \textit{Chidamber and Kemmerer} set of object-oriented metrics \cite{CHKE94}. The version of the tool that we used for this experiment can be found in GitHub~\cite{CKJM}.

  \item \textbf{JDepend} \textit{(jdep)}~\cite{JDEPEND}; a tool that analyses {\sc jar} files that contain compiled Java classes and calculates a series of design metrics.

  \item \textbf{CLMT} \textit{(clmt)}~\cite{SGKL09}; which stands for \textit{cross-language metric tool} and analyses the source code of several languages in order to calculate a series of size metrics.
\end{itemize}

@@

Table \ref{tbl:oss-size-metrics} presents several size metrics for our dataset. It includes 11,365 projects, with more than 74 million lines of code and almost 33 million unique measurements. Table~\ref{tbl:class-selected-metrics}, \ref{tbl:method-selected-metrics}, \ref{tbl:package-selected-metrics}, and \ref{tbl:size-selected-metrics} present the key metrics that are calculated and stored in the dataset. Each tool focuses on a different aspect of a software system; \textit{ckjm} focuses on object-oriented design metrics, \textit{jdepend} on package design, while \textit{clmt} focuses on program size metrics. Several metrics are calculated by more than one tool, like \textit{cyclomatic complexity}, which is calculated by both \textit{clmt} and \textit{ckjm}. Both calculations are available in the database.

% dimitro@bkarak: enrich the above with refs. The ref section is poor. e.g. [...] focuses on object-oriented design metrics *citation here* [...].

A series of newly introduced metrics were also stored in the database. Such metrics can count the usage of specific {\sc dsl} application libraries in a software project and they are included in the Table~\ref{tbl:size-selected-metrics}. The methodology to identify and calculate the {\sc dsl} metrics was the following: A set of standard {\sc dsl} application libraries was identified, and the source was scanned for specific \textit{import} statements (e.g. \textit{java.util.regex}). These statements indicated that the standard package that implements regular expressions was used, thus regular expressions were used in the project. Build files or other resources that may contain {\sc dsl}s were not included. One final assumption was also made; if {\sc xp}ath or {\sc xslt} were found in the source code, then the project would be marked as a project that utilizes {\sc xml}. This is because both languages are used for query and transformations on {\sc xml} {\sc dom} trees. Table \ref{tbl:dsl-list} lists the selected {\sc dsl}s application libraries. Note that we focused on libraries that were included as part of the Java {\sc sdk}.

@@

In Section~\ref{sec:dsl}, we describe an experiment based on these metrics. The experiment measures the popularity of {\sc dsl} usage for the dataset and therefore for the Java ecosystem in general.
